\section*{Data and code availability statement}

Both datasets that were used in the current study are publicly available. The LEMON dataset \citep{Babayan2019} can be accessed at: \url{https://fcon_1000.projects.nitrc.org/indi/retro/MPI_LEMON.html}. The RIFT dataset \citep{Spaak2024_dataset} is available for download from: \url{https://doi.org/10.34973/8dx5-7e51}. All analysis scripts are available at: \url{https://github.com/ctrltz/crosstalk-evaluation}. In addition, we provide two Python packages for (1) simulation of M/EEG activity (MEEGsim; \url{https://meegsim.readthedocs.io}) and (2) evaluation of spatial filters using CTF (ROIextract; \url{https://roiextract.readthedocs.io/}).

\section*{Ethics statement}

No data was collected in this study. For the LEMON dataset, data collection was conducted in accordance with the Declaration of Helsinki, and the study protocol was approved by the ethics committee of the medical faculty at the University of Leipzig (reference number 154/13-ff). For the RIFT dataset, the data collection was approved by the local ethics committee (CMO Arnhem-Nijmegen, Radboud University Medical Center) under the general ethical approval for the Donders Centre for Cognitive Neuroimaging ('Imaging Human Cognition', CMO 2014/288).

\section*{CREDiT authorship contribution statement}

\textbf{Nikolai Kapralov:} Conceptualization, Data curation, Formal analysis, Investigation, Methodology, Project administration, Software, Validation, Visualization, Writing – original draft, Writing – review \& editing. \textbf{Alina Studenova:} Conceptualization, Investigation, Formal analysis, Software, Writing – review \& editing. \textbf{Rubén Eguinoa:} Formal analysis, Software, Writing – review \& editing. \textbf{Guido Nolte:} Methodology, Writing – review \& editing. \textbf{Stefan Haufe:} Supervision, Writing – review \& editing. \textbf{Arno Villringer:} Funding acquisition, Supervision, Writing – review \& editing. \textbf{Vadim Nikulin:} Conceptualization, Investigation, Methodology, Supervision, Writing – review \& editing.

\section*{Declaration of competing interests}

The authors declare no competing interests.

\section*{Acknowledgments}

This work was supported by the Max Planck Society and, in part, by the cooperation project between the Max Planck Society and the Fraunhofer Gesellschaft (grant: project NEUROHUM). The work of Rubén Eguinoa was supported by the Spanish Ministry of Research and Innovation under Grant PID2020-118829RB-I00. Data were provided (in part) by the Radboud University, Nijmegen, The Netherlands.
